Generating Hubbard Model Solutions from Anderson Impurity Model Solutions

### 2021-08-07 notes.tex
cosmetic changes and elaborations

\section{Computing single particle Greens functions and related quantities}

@@

	\label{fig:}

@@

@@

	\ket{i} &= \frac{1}{\sqrt N}\sum_{\vec k=\vec k_1}^{\vec k_N} e^{i \vec{k}\cdot\vec{r_i}}\ket{\vec k}\\

@@

@@

The matrix is of course written in momentum space, and diagonal. Using Dyson's equation for the self-energy $\Sigma = G_{0}^{-1} - G^{-1}$ (where the matrix $G_{0,H}^{-1} (\omega) = \sum_{\vec k=\vec k_1}^{\vec k_N} (\omega -\epsilon_{k})\ket{\vec k}\bra{\vec k}=\sum_{\vec k=\vec k_1}^{\vec k_N} (\omega +t^{H}\xi_{k})\ket{\vec k}\bra{\vec k} $ is the inverse Greens function for the appropriate non-interacting tight-binding system), we obtain the self-energy matrix as

@@

&=&\sum_{\vec k=\vec k_1}^{\vec k_N} (\omega - g_{0} + (t^{H} - g_{1})\xi_{\vec{k}})\ket{\vec k}\bra{\vec k} = \sum_{\vec k=\vec k_1}^{\vec k_N} \Sigma_{H} (\vec{k},\omega) \ket{\vec k}\bra{\vec k}~,\nonumber\\
\Sigma_{H} (\vec{k},\omega) &=& \omega - g_{0} + (t^{H} - g_{1})\xi_{\vec{k}}~.

@@

Thus, we can use eqs.\eqref{2dsquaretb},\eqref{local_gf} and \eqref{nn_gf} to obtain the full self-energy $\Sigma (\vec{k},\omega)$ for the Hubbard model on the 2D square lattice.

@@

	G_H(\omega) = \sum_{\vec k=\vec k_1}^{\vec k_N}\left(g_0 + g_1 \xi_{\vec k}\right)^{-1}\ket{\vec k}\bra{\vec k} = \begin{pmatrix} \left(g_0 + g_1\xi_{\vec k_1}\right)^{-1} &&& \\
	& \left(g_0 + g_1\xi_{\vec k_2}\right)^{-1} && \\
	&& \left(g_0 + g_1\xi_{\vec k_3}\right)^{-1} & \\

@@

The real-space Greens function, $G_{H}(\vec{r},\omega)$, can now be obtained by Fourier transforming $G_H(\vec k, \omega)$ 

@@

	G_H(\vec r, \omega) = \frac{1}{\sqrt N}\sum_{\vec k} e^{i\vec{k}.\vec{r}}G_H(\vec k, \omega) &= \frac{1}{\sqrt N}\sum_{\vec k}e^{i\vec{k}.\vec{r}}\left(g_0 + g_1 \xi_{\vec k}\right)^{-1}\\
												     &= \frac{\sqrt N}{2}\sum_{\vec k}e^{i\vec{k}.\vec{r}}\left\{\left[G_{D}^{-1}(\omega)\right]_{00\sigma} + \frac{1}{w}\left[G_{D}^{-1}(\omega)\right]_{01\sigma}\xi_{\vec k}\right\}^{-1}

@@

Thus, we can now compute the real space local Green's function $G_H(\vec r = 0, \omega)$ by summing over the momenta:

@@

	G_H(\vec r = 0, \omega) = \frac{1}{\sqrt N}\sum_{\vec k} G_H(\vec k, \omega) = \frac{1}{\sqrt N}\sum_{\vec k}\left(g_0 + g_1 \xi_{\vec k}\right)^{-1}

@@

The simplest test involves calculating \(G_H(\vec r = 0, \omega)\) for the Hubbard dimer \(\left(N = 2, w = 1 \right)\). The discrete set of momenta are $0, \frac{\pi}{a}$, so

@@

	\xi_{\vec k} = 1, -1

@@

Therefore, the Greens function at $\vec r=0$ for the Hubbard dimer can be expressed as
	G_H(\vec r = 0, \omega) = \frac{1}{2}\left[\frac{1}{g_0 + g_1} + \frac{1}{g_0 - g_1}\right]  = \frac{g_0}{g_0^2 - g_1^2}
$g_{0,1}$ are expressed in terms of the elements of the inverse Greens function matrix of the Hubbard dimer. The inverse of a $2\times 2$ matrix is
	G_D^{-1} = \frac{1}{\text{Det }G_D}\begin{pmatrix} \left(G_D\right)_{00} &  -\left(G_D\right)_{01} \\ -\left(G_D\right)_{01} & \left(G_D\right)_{00}\end{pmatrix} \implies \begin{cases}
		g_0\equiv\frac{2}{N} (G^{-1}_{D})_{00} = \frac{2}{N} \frac{1}{\text{Det }G_D}\left(G_D\right)_{00} \\
		\\
		g_1\equiv\frac{2}{Nw} (G^{-1}_{D})_{00} = -\frac{2}{Nw}\frac{1}{\text{Det }G_D}\left(G_D\right)_{01} \\
	\end{cases}
Using the expressions for $(G_{D})_{00}$ and $(G_{D})_{01}$ shown in the appendix, we can obtain $g_{0}$ and $g_{1}$ very generally. For the Hubbard dimer, we substitute $N=2, w=1$ into the expressions for $g_{0}$ and $g_{1}$ to obtain
	G_H(\vec r = 0, \omega) = \text{Det }G_D\frac{\left(G_D\right)_{00}}{\left(G_D\right)_{00}^2 - \left(G_D\right)_{01}^2} = \text{Det }G_D\frac{\left(G_D\right)_{00}}{\text{Det }G_D} = \left(G_D\right)_{00}
This is the correct result, because the full Hubbard model in this case was simply the Hubbard dimer, so \(G_H(\vec r = 0, \omega)\) should just be \(\left(G_D\right)_{00}\). Although this is a trivial example (we matched the Hubbard dimer with another Hubbard dimer), this atleast tells us that this algorithm for calculating the real space local Greens function from the inverse Greens function matrix elements is sensible.

@@

The site local spectral function can now be calculated:

>>>>

\section{Formal expressions for single particle Greens functions and other related many-body quantites}
\subsection{Expressing matrix elements of the inverse single particle Greens function in terms of Hubbard dimer counterparts}

@@

@@

\subsection{Constructing full Greens function matrix from the inverse matrix via Fourier transformation to \(k-\)space}

@@

\begin{flalign*}
	\ket{i} &= \frac{1}{\sqrt N}\sum_{\vec k=\vec k_1}^{\vec k_N} e^{i \vec{k}\cdot\vec{r_i}}\ket{\vec k} \Leftrightarrow \ket{k} = \frac{1}{\sqrt N}\sum_{\vec i} e^{-i \vec{k}\cdot\vec{r_i}}\ket{\vec r_i} \\

@@

\end{flalign*}

@@

The matrix entries are written in momentum space, and the matrix is diagonal in this basis.
\subsection{Calculation of self energy matrix from the Dyson equation}
Using Dyson's equation for the self-energy $\Sigma = G_{0}^{-1} - G^{-1}$ (where the matrix $G_{0,H}^{-1} (\omega) = \sum_{\vec k=\vec k_1}^{\vec k_N} (\omega -\epsilon_{k})\ket{\vec k}\bra{\vec k}=\sum_{\vec k=\vec k_1}^{\vec k_N} (\omega +t^{H}\xi_{k})\ket{\vec k}\bra{\vec k} $ is the inverse Greens function for the appropriate non-interacting tight-binding system), we obtain the self-energy matrix as

@@

&=&\sum_{\vec k=\vec k_1}^{\vec k_N} (\omega - g_{0} + (t^{H} - g_{1})\xi_{\vec{k}})\ket{\vec k}\bra{\vec k}

@@

This is the full self energy matrix. One can get both real space and momentum space self energies by taking appropriate matrix elements. For instance, the self-energy in momentum space is
	\label{selfenergy}
	\Sigma_{H} (\vec{k},\omega) \equiv \bra{\vec k} \Sigma_H (\omega) \ket{\vec k} = \omega - g_{0} + (t^{H} - g_{1})\xi_{\vec{k}}~.
Thus, we can use eqs.\eqref{2dsquaretb}, \eqref{local_gf} and \eqref{nn_gf} to obtain the full self-energy $\Sigma (\vec{k},\omega)$ for the Hubbard model on the 2D square lattice.
\subsection{Obtaining real space Greens functions from the full matrix}

@@

	\label{k-space-gf}
	G_H(\omega) = \sum_{\vec k=\vec k_1}^{\vec k_N}\left(g_0 + g_1 \xi_{\vec k}\right)^{-1}\ket{\vec k}\bra{\vec k} = \begin{pmatrix} \left(g_0 + g_1\xi_{\vec k_1}\right)^{-1} &&&&\\
	& \left(g_0 + g_1\xi_{\vec k_2}\right)^{-1} &&&\\
	&& \left(g_0 + g_1\xi_{\vec k_3}\right)^{-1} &&\\
		...&&&...&\\

@@

The real-space Greens functions can now be obtained by taking matrix elements of the full Greens function matrix in eq.~\ref{k-space-gf}:

@@

	G_H(\vec r_i, \vec  r_j, \omega) \equiv \bra{\vec r_i} G_H(\omega) \ket{\vec r_j} = \sum_{\vec k} G_H(\vec k, \omega)  \left<\vec r_i | \vec k\right>\left<\vec k | \vec r_j\right>

@@

From the definition of the Fourier transforms mentioned previously, we can write

@@

	\left<\vec r_i | \vec k\right> = \frac{1}{\sqrt N}\sum_{\vec r} e^{- \vec{k}\cdot\vec{r}} \left<\vec r_i | \vec r \right> = \frac{1}{\sqrt N}e^{- \vec{k}\cdot\vec{r_i}}
Substituting this gives
	G_H(\vec r_i, \vec  r_j, \omega) = \frac{1}{N}\sum_{\vec k} G_H(\vec k, \omega)  e^{\vec{k}\cdot\left(\vec r_j - \vec{r_i}\right)} \equiv G_H(\vec r_j - \vec  r_i, \omega)
The final expression brings out the fact that the Greens function for excitations between \(\vec r_i\) and \(\vec r_j\) depends only on the relative position \(\vec r_j - \vec r_i\), not on their absolute positions. This is expected on grounds of translational invariance. By choosing \(\vec r_i = \vec r_j\), we can obtain the real space \textit{local} Greens function:
	\left(G_H\right)_{ii}(\omega) \equiv G_H(\vec r_j - \vec  r_i = 0, \omega) = \frac{1}{N}\sum_{\vec k} G_H(\vec k, \omega) = \frac{1}{N}\sum_{\vec k}\left(g_0 + g_1 \xi_{\vec k}\right)^{-1}

@@

\subsection{Obtaining \(g_0\) and \(g_1\)}
These two quantities are expressed in terms of matrix elements of the inverse Greens function \(G_D^{-1}\) of the Hubbard dimer. Since \(G_D\) is a \(2\times 2\) matrix, it is easy to invert:

@@

	G_D &= \begin{pmatrix} \left(G_D\right)_{00} &  \left(G_D\right)_{01} \\ \left(G_D\right)_{01} & \left(G_D\right)_{00}\end{pmatrix} \\
	\implies G_D^{-1} &= \frac{1}{\text{Det }G_D}\begin{pmatrix} \left(G_D\right)_{00} &  -\left(G_D\right)_{01} \\ -\left(G_D\right)_{01} & \left(G_D\right)_{00}\end{pmatrix}\\
		 \implies g_0 &\equiv\frac{2}{N} (G^{-1}_{D})_{00} = \frac{2}{N} \frac{1}{\text{Det }G_D}\left(G_D\right)_{00},\\
		 g_1 &\equiv\frac{2}{Nw} (G^{-1}_{D})_{00} = -\frac{2}{Nw}\frac{1}{\text{Det }G_D}\left(G_D\right)_{01} 

@@

Using the expressions for $(G_{D})_{00}$ and $(G_{D})_{01}$ shown in the appendix, we can obtain $g_{0}$ and $g_{1}$ very generally.

\section{Analytic consistency checks}
\subsection{Single-particle Greens function and self energy for the Hubbard dimer}
The simplest test involves calculating \(G_H(\vec r = 0, \omega)\) for the Hubbard dimer \(\left(N = 2, w = 1 \right)\). The discrete set of momenta are \(\left\{ \vec k_n \right\} = 0, \frac{\pi}{a}\), so
\begin{equation}\begin{aligned}
	\left\{\xi_{\vec k}\right\} = 1, -1
\end{aligned}\end{equation}
Therefore, the Greens function at $\vec r=0$ for the Hubbard dimer can be expressed as
\begin{equation}\begin{aligned}
	G_H(\vec r = 0, \omega) = \frac{1}{2}\left[\frac{1}{g_0 + g_1} + \frac{1}{g_0 - g_1}\right]  = \frac{g_0}{g_0^2 - g_1^2}
\end{aligned}\end{equation}
For the Hubbard dimer, we substitute $N=2, w=1$ into the expressions for $g_{0}$ and $g_{1}$ to obtain
\begin{equation}\begin{aligned}
	&g_0 = \frac{1}{\text{Det }G_D}\left( G_D \right)_{00}, &g_1 = -\frac{1}{\text{Det }G_D}\left( G_D \right)_{01}
\end{aligned}\end{equation}
The Greens function at the zeroth site then becomes
\begin{equation}\begin{aligned}
	G_H(\vec r = 0, \omega) = \text{Det }G_D\frac{\left(G_D\right)_{00}}{\left(G_D\right)_{00}^2 - \left(G_D\right)_{01}^2} = \text{Det }G_D\frac{\left(G_D\right)_{00}}{\text{Det }G_D} = \left(G_D\right)_{00}
\end{aligned}\end{equation}
This is the expected result. Because the full Hubbard model in this case was simply the Hubbard dimer, \(G_H(\vec r = 0, \omega)\) should just be \(\left(G_D\right)_{00}\). Although this is a trivial example (we matched the Hubbard dimer with another Hubbard dimer), this atleast tells us that this algorithm for calculating the real space local Greens function from the inverse Greens function matrix elements is sensible.

\subsection{On the Bethe lattice}

@@

In the atomic limit, the singly-occupied state has zero energy:
	E_1(t=0) = \bra{1,0}\left(U \tau_{0 \uparrow}\tau_{0 \downarrow} + U \tau_{1 \uparrow}\tau_{1 \downarrow}\right) \ket{1,0} = 0
We can write the atomic limit Greens function in terms of this energy and the self energy:
	G_{D,00}^\uparrow(\omega, T \to 0) \bigg\vert_\text{atomic} = \frac{1}{\omega - E_1(t=0) - \Sigma(t=0)} = \frac{1}{\omega - 0 -\frac{U^2}{4\omega}}
The self energy in the atomic limit can be read off as 
	\label{dimer_selfenergy}
\Sigma(t=0) = \frac{U^2}{4\omega}
The site local spectral function can also be calculated from the local Greens function:
